\documentclass[10pt]{article}
\input{common}
\title{RevIN Experiment 1: Normalization Components}
\author{}
\date{}
\begin{document}\maketitle
\section{Theoretical overview and goal}
Global standardization applies training-wide statistics.  Reversible instance
normalization instead uses the current look-back:
\[
 \tilde x=\alpha\frac{x-\mu_x}{\sigma_x}+\beta,
 \qquad
 \hat y=\sigma_x\frac{f_\theta(\tilde x)-\beta}{\alpha}+\mu_x.
\]
Non-affine instance normalization fixes $(\alpha,\beta)=(1,0)$.  It removes
window-specific offset and scale, which can reduce temporal and cross-user
shift, but it also discards potentially predictive level information.

\goal{Isolate the contribution of global normalization, reversible per-instance
normalization, and RevIN's learnable affine parameters under the same
data-space training loss.}

\section{Implementation details}
The primary MSE-trained comparison is \code{none\_mse},
\code{standard\_mse}, \code{instance\_mse}, and \code{revin\_mse}.  Standard
statistics are fitted on training data only.  Instance and RevIN statistics are
computed from each look-back and inverted after forecasting; all reported MSEs
are in original data space.

The full sweep uses ETTh1 (target OT only), Electricity, Traffic, Solar, Weather, and
Exchange Rate; six $(L,H)$ settings; PatchTST; three seeds; batch
size 256; learning rate $10^{-5}$; and exactly 10,000 optimizer steps.  The long-horizon setting
is $(336,720)$. Validation and logging run every 1,000 optimizer steps, while
evaluation windows use a horizon-sized stride. Training windows are sampled at
random over users and dates. Test mode uses Electricity/Solar, two settings,
PatchTST, seeds 1--2, and 2,000 optimizer steps.

\textbf{Implementation:} \code{src/utils/models.py}\\
\textbf{Launcher:} \code{revin.slurm}

Run isolated \code{test} mode first. The \code{small} profile then reproduces
the original Traffic/Electricity/Solar PatchTST study on four settings. The
\code{full} profile shares its output root, reuses completed small seeds, and
adds all datasets, two comparison settings, standard--nMSE, last-value, and
arcsinh runs. The \code{ultra} profile adds DLinear.
One table is produced per model and test split,
with seed mean $\pm$ sample standard deviation and an explicit row multiplier
$\times10^m$. JSON summaries also retain the variance of per-point losses.

\section{Interpretation}
The none--standard contrast measures global conditioning.  Standard--instance
measures the value of sample-specific alignment.  Instance--RevIN isolates the
affine parameters; it should not be described as the effect of instance
normalization itself.
\end{document}
