\documentclass[10pt]{article}
\input{common}
\title{RevIN Experiment 2: Normalized Backpropagation}
\author{}
\date{}
\begin{document}\maketitle
\section{Theoretical overview and goal}
With instance statistics $(\mu_x,\sigma_x)$, data-space MSE is
$\lVert\hat y-y\rVert^2$, whereas normalized backpropagation minimizes
\[
 \operatorname{nMSE}(\hat y,y\mid x)=
 H^{-1}\sum_{h=1}^{H}
 \left(\frac{\hat y_h-y_h}{\sigma_x+\varepsilon}\right)^2.
\]
The latter gives comparable gradient weight to low- and high-scale windows.
Predictions are still denormalized and evaluated with ordinary data-space MSE,
so only optimization weighting changes.

\goal{Test whether scale-balanced gradients improve generalization independently
of the forward normalization layer, and whether RevIN's affine parameters alter
that effect.}

\section{Implementation details}
Two paired contrasts are run:
\begin{center}
\begin{tabular}{lll}
\toprule
Forward transform & Data-space loss & Normalized loss \\
\midrule
Non-affine instance & \code{instance\_mse} & \code{instance\_nmse} \\
Affine RevIN & \code{revin\_mse} & \code{revin\_nmse} \\
\bottomrule
\end{tabular}
\end{center}
Architecture, seed, batch composition, learning rate, and epoch budget are held
fixed within every pair.  The project supports MSE, MAE, nMSE, nMAE, and
relative MSE, but this paper comparison focuses on MSE versus nMSE.

Before splitting, the loader automatically applies the sibling dataset
\code{config.json}: top-level and \code{revin} user exclusions are combined with
run-specific additions and recorded in each seed directory. This curated list
does not replace the separate constant-window audit required by nMSE.

\textbf{Loss implementation:} \code{src/utils/pipeline.py}\\
\textbf{Experiment orchestration:} \code{src/scripts/experiment.py}

The small sweep covers Traffic, Electricity, and Solar, four settings,
PatchTST, and the six core methods. The full sweep expands to six datasets, six
settings, and all loss/normalization variants while retaining PatchTST; ultra
adds DLinear. All publication profiles use three seeds and
exactly 10,000 optimizer steps, with validation every 1,000 steps and
horizon-strided evaluation. Final
tables always use test MSE in data space, show sample standard deviations, and
state the row multiplier. Macro summaries give each dataset--setting pair equal
weight and include a test-selected standard/instance oracle only as an
optimistic reference.

\section{Interpretation}
An nMSE-trained model can win on data-space MSE because balanced gradients learn
a mapping that generalizes across heterogeneous scales.  Training curves from
MSE and nMSE have different units and should not be compared by absolute height;
selection should use the common validation metric.
\end{document}
