\documentclass[10pt]{article}
\input{common}
\title{RevIN Experiment 3: Centering and Transform Variants}
\author{}
\date{}
\begin{document}\maketitle
\section{Theoretical overview and goal}
Mean centering assumes the look-back average is the appropriate location anchor
for the future.  Last-value centering instead predicts changes around the most
recent observation, which may better suit local trends.  The inverse hyperbolic
sine transform
\[
 z=\operatorname{asinh}(x)
\]
is approximately linear near zero and logarithmic for large magnitudes, giving
a reversible compression of heavy-tailed values before instance normalization.

\goal{Test two lightweight changes to RevIN's inductive bias: a more local
location anchor and a nonlinear magnitude compression as appendix ablations.}

\section{Implementation details}
The variants use non-affine instance normalization and normalized MSE:
\begin{center}
\begin{tabular}{lll}
\toprule
Method & Center & Transform \\
\midrule
\code{instance\_nmse} & Look-back mean & Identity \\
\code{revin\_last\_nmse} & Last observation & Identity \\
\code{revin\_arcsinh\_nmse} & Look-back mean & asinh \\
\bottomrule
\end{tabular}
\end{center}
Every transformation is inverted before data-space evaluation.  The same six
datasets, seven settings, DLinear/PatchTST models, five seeds, optimizer, batch
size, epoch budget, and validation frequency are used.  Configuration mapping
is defined by \code{method\_args} in
\code{src/slurm/run\_revin\_experiment.slurm}; layer behavior is in
\code{src/utils/models.py}.

Tables are separate by model and test split.  They report ordinary MSE as mean
$\pm$ sample standard deviation with explicit $\times10^m$ row scaling.  The
baseline for both contrasts is \code{instance\_nmse}.

\section{Interpretation}
Last centering tests sensitivity to local level evolution; asinh tests
sensitivity to tails and multiplicative scale.  A gain is evidence for that
specific preprocessing bias, not a general solution to conditional shift.
Results should be examined across datasets because either transform may remove
useful absolute-level information.
\end{document}
